Rule-Based Preordering on Multiple Syntactic Levels \\in Statistical Machine Translation

### 2014-09-25 Paper/paper.tex
.

\documentclass[a4paper]{article}
\usepackage{iwslt14,amssymb,amsmath,epsfig,multirow,spreadtab,array}
\usepackage{tikz}
\usetikzlibrary{arrows}
\usetikzlibrary{trees}
\usetikzlibrary{shapes}
\usetikzlibrary{chains}
\usetikzlibrary{calc}
\usetikzlibrary{positioning}
\usepackage[T1]{fontenc}
\usepackage[latin1,utf8x]{inputenc}
\usepackage[encapsulated]{CJK}
\newcommand{\cntext}[1]{\begin{CJK}{UTF8}{gbsn}#1\end{CJK}}
\setcounter{page}{1}
\sloppy		% better line breaks
%\ninept
%SM below a registered trademark definition
\def\reg{{\rm\ooalign{\hfil
     \raise.07ex\hbox{\scriptsize R}\hfil\crcr\mathhexbox20D}}}

@@

\title{Rule-Based Preordering on
Multiple Syntactic Levels in
Statistical Machine Translation}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Please make sure to keep technical paper submissions anonymous  !
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\name{Firstname Lastname}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% If multiple authors, uncomment and edit the lines shown below.       %%
%% Note that each line must be emphasized {\em } by itself.             %%
%% (by Stephen Martucci, author of spconf.sty).                         %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \makeatletter
 \def\name#1{\gdef\@name{#1\\}}
 \makeatother
 \name{{\em Ge Wu, Yuqi Zhang}}
%%%%%%%%%%%%%%% End of required multiple authors changes %%%%%%%%%%%%%%%%%

\address{Institute for Anthropomatics \\
Karlsruhe Institute of Technology, Germany \\
{\small \tt utcur@student.kit.edu, yuqi.zhang@kit.edu}
}
%
\begin{document}
\maketitle
%
\begin{abstract}
We propose a novel data-driven rule-based preordering approach, which uses the information of syntax tree and word alignment to reorder the words in source sentences before decoding in a phrase-based SMT system between English and Chinese. The preordering algorithm extracts reordering patterns from multiple levels of the syntax trees of the training data and applies the rules on the source sentences in a similar way. We've conducted experiments in English-to-Chinese and Chinese-to-English translation directions. Our results show that the approach has led to improved translation quality both when it was applied separately on the baseline or when it was combined with some other reordering approaches. As our reordering approach was used alone, it showed an improvement of $1.61$ in BLEU score in the English-to-Chinese translation direction and an improvement of $2.16$ in BLEU score in the Chinese-to-English translation direction, in comparison with the baseline, which used no word reordering. As our preordering approach were combined with the short rule \cite{short}, long rule \cite{long} and tree rule \cite{tree} based preordering approaches, it showed further improvements of up to $0.43$ in BLEU score in the English-to-Chinese translation direction and further improvements of up to $0.3$ in BLEU score in the Chinese-to-English translation direction. Through the translations that used our preordering approach, we've also found improvements in syntactic structure.
\end{abstract}
%
\section{Introduction}

Word order is a general issue when we want to translate text from one language to the other. Different languages normally have different word orders and the difference could be very huge. Among all the languages, Chinese is one language which is very different from English, because they belong to different language families and have long period of separately development. Both languages have a Subject-Verb-Object order, but they also have a lot of differences in word order. Especially sentences in both languages can sometimes have completely different syntactic structures. The differences may involve long-distance or unstructured position changes.

@@

\section{Related Work}

@@

In a phrase-based SMT system, there are several possibilities to change the word orders. Words can be reordered during the decoding phase by setting a window, which allows the decoder to choose the next word for translation. Reordering could also be influenced by the language model, because the language model give probability of how a certain word is likely to follow. Different language model may give different probability, which further influences the decision made by log-linear model. Other ways to change the word orders include using distance based reordering models or lexicalized reordering models \cite{tillmann2004, koehn2005}. The lexicalized reordering model reorders the phrases by using information of how the neighboring phrases change orientations.

@@

However, these approaches which are bases on POS tag sequences or syntax trees are mostly designed for languages like German and are not especially adapted for languages like Chinese. As Chinese has very different word orders, a reordering approach, which can further explore the syntactic structure of Chinese and utilize this information for reordering, is desirable.

@@

\section{Motivation}

@@

Some typical problems of word orders between English and Chinese that we've found in the data are as follows:

\begin{itemize}
\item \emph{Pre-modifier instead of post-modifier}\\ In Chinese people tend to use pre-modifier rather than post-modifier. This involves the position change of adverbials, relative clauses and preposition phrases during translation.
\item \emph{Construction of questions}\\
The two languages have very different ways to construct questions, which raises word order problems for translating questions.
\item \emph{Special sentence constructions}\\ For example, \emph{B\^{a}}-construction in Chinese and sentence constructions such as \emph{there be} and inverted negative sentences in English don't have correspondence in the other language in general.
\item \emph{Long distance word position change}\\
Word reordering between English and Chinese often involves word shift cross long distance.
\end{itemize}

From previous work of others, the following types of reordering rules have been researched:

\begin{itemize}
\item \emph{Short rules} \cite{short} \\
Short rules are extracted based on the sequences of adjacent words or their POS tags in sentences from training data. Sequences of adjacent words or POS tags are observed, rules are then extracted if the same reordering patterns appear frequently. Following is a reordering pattern of a short rule:\medskip\\
\texttt{WRB MD DT -> DT WRB MD}
\item \emph{Long rules} \cite{long}\\
Long rules are specially designed to improve the long distance word reordering for language pairs like English and German. The rules are based on POS tags of the text, and use a placeholder ($*$) to represent one or more words. Following is a reordering pattern of a long rule:\medskip\\
\texttt{NN $*$ MD -> $*$ MD NN}

\item \emph{Tree rules} \cite{tree}\\
While short rules and long rules are based on the flat structure of sentences, tree rules reorder sentences by using information from sentences' syntactic structure. The syntax tree and word alignment of the training corpus are used to train the reordering rules. The tree rules reorder the words both on the word level and on the constituent level. Following is a reordering pattern of a tree rule:\medskip\\
\texttt{NP ( ADJP JJ NN ) -> JJ NN ADJP}%\medskip\\
%The parenthesis in the example represents the hierarchy in the syntax tree. The left side of the rule corresponds a tree with the root labeled with \emph{NP} and three children, each labeled with \emph{ADJP}, \emph{JJ} and \emph{NN}. 
\end{itemize}

However, not all these reordering problems of translating Chinese can be solved well by using these reordering rules based either on POS tag sequence or single level of syntax tree. In order to improve the reordering between English and Chinese, we need reordering method that can handle more complicated, unstructural word order change. Inspired by the ideas of reordering on syntax tree and hierarchical phrases, we created the MLT reordering, which reorders words based on multiple syntactic levels and can handle long distance word position change and complicated word position change very well.

\section{Multi-Level-Tree Reordering}

Our preordering method is based on reordering rules. Reordering rules show how sentences should be reordered in source language before translation. In our system, the rules are generated by using the word alignment, and syntax tree, all of which are calculated based on the training data. After reordering rules are applied to the source sentences, word lattices are generated. A word lattice contains all the reordering possibilities of a source sentence and is further passed to the decoder for translation. The preordering system could be illustrated as figure~\ref{prereordering}.

\begin{figure*}
\centering
\input{prereordering.tikz}
\caption{Illustration of preordering system}
\label{prereordering}
\end{figure*}

@@

Reordering patterns are based on multiple levels of the syntax tree. Figure~\ref{unstructured} illustrates how the reordering patterns are detected from the syntax tree. In the example, the detection starts from the root node, go downwards three levels and use the nodes in these levels to detect the reordering pattern. These nodes that are used for detecting the reordering pattern are colored gray and have an italic font. The leaf nodes in the syntax tree indicate words in the sentence, which has rectangle shape with angular corners.

According to the alignment information with the corresponding Chinese translation, the node labeled with \emph{NP} is moved forward to the first place in the translation and the node labeled with \emph{IN of} is moved forward to the second place in translation. So from the root node with a search depth $3$, the following reordering pattern can be found:\bigskip

\noindent \texttt{NP ( \textbf{CD}$_0$ NP ( NP ( \textbf{JJ}$_1$ \textbf{NNS}$_2$ ) PP ( \textbf{IN}$_3$ \textbf{NP}$_4$ ) ) ) -> NP IN CD JJ NNS}\medskip \\
\texttt{-> 4 3 0 1 2 (alternative with index)}\bigskip

The POS tags in bold corresponds to words in the leaf nodes of the syntax tree, which are parts that are actually reordered and have a darker gray color in the figure. The parentheses in the reordering pattern indicate the corresponding hierarchies in the syntax tree. The reordering can be alternatively represented with index, which is the actual internal representation to avoid ambiguity caused by POS tags with the same name and we use this representation throughout this paper.

\begin{figure}
\centering
\input{unstructured.tikz}
\caption{Detection of reordering pattern from multiple syntactic levels}
\label{unstructured}
\end{figure}

@@

In order to find as much information for reordering as possible. The algorithm of rule extraction detects the reordering patterns from all nodes in the syntax tree and it goes downwards for any number of hierarchies, until it reaches the lowest hierarchy in the subtrees.

@@

\begin{figure}
\centering
\input{extract.tikz}
\caption{A phrase with its syntax tree and word alignment for rule extraction}
\label{extract}
\end{figure}

Figure~\ref{extract} shows a phrase to be translated, together with its syntax tree and word alignment of parallel text. The leaf nodes are underlined and indexed according to the order they appear in the left side of the patterns. The index sequences on the right side of the patterns indicate the reordering. In this example, we can find the following reordering patterns:\bigskip

\noindent From node $1$:\\
\texttt{NP ( \textbf{NP} \textbf{PP} ) -> 1 0} \hfill [1 level]\hphantom{s}\\
\texttt{NP ( NP ( \textbf{JJ} \textbf{NNS} ) PP ( \textbf{IN} \textbf{NP} ) ) -> 3 2 0 1} \hfill [2 levels]\\
\texttt{NP ( NP ( \textbf{JJ} \textbf{NNS} ) PP ( \textbf{IN} NP ( \textbf{JJ} \textbf{NNS} ) ) ) -> 3 4 2 0 1} \hfill [3 levels]\medskip

\noindent From node $3$:\\
\texttt{PP ( \textbf{IN} \textbf{NP} ) -> 1 0} \hfill [1 level]\hphantom{s}\\
\texttt{PP ( \textbf{IN} NP ( \textbf{JJ} \textbf{NNS} ) ) -> 1 2 0} \hfill [2 levels]\bigskip

@@

The syntax tree is traversed by DFS as the same in rule extraction. But from the root of each subtree, it's scanned with depth limit from its maximal levels, i.e. its depth, to $1$. As it turns out, any rule can be applied for a subtree at some level, a new path for this reordering will be added to the word lattice for decoding. As long as rules can be applied on a subtree for a certain depth, the rules are applied and the search for rule application on this subtree stops, and the search on the next subtree continues.

The reason for this is to prevent duplicate reorderings due to application of nested rules, which has overlapped effect with each other. These rules are normally patterns that are generated on the same subtree, but with different number of levels, which has different generalization effect on the same range of words in the text. For example, the following patterns can be detected from the syntax tree in figure~\ref{extract}:\bigskip

\noindent \texttt{PP ( \textbf{IN} \textbf{NP} ) -> 1 0}\\
\texttt{PP ( \textbf{IN} NP ( \textbf{JJ} \textbf{NNS} ) ) -> 1 2 0}\bigskip

Both patterns are detected from the same node, but the second pattern is detected by retrieving the nodes one level deeper and it's more concrete. So the first pattern can be seen as a generalization of the second pattern. Whenever a rule of the second pattern can be applied, a rule of the first pattern can be applied too. Because subtrees are checked from the highest number of levels in rule application, the more concrete rule is applied first. Because the more concrete rule fits the detected pattern better and contains more details of reordering, so it may be more suitable for rule application. In this example, the second rule is applied rather than the first rule.

@@

In order to further explore the probability of improvement, the MLT reordering rules can be combined with other types of reordering rules to achieve further improvements. This is done by training the different types of rules separately and applying them on the monotone path of the sentence independently. They generate different paths in the word lattice.

\section{Experiments}

We've conducted experiments on both English-to-Chinese and Chinese-to-English translation directions to get a better overview of the MLT reordering's effect. In the experiments, the MLT reordering rules are also combined with other reordering rules that are introduced before, in order to show the improvement achieved by our approach.

In order to evaluate the potential of word reordering, we also used oracle reordering in the experiments. Oracle reordering is considered to be an optimally reordered sentence as input to the SMT system and do not allow additional reordering during decoding.\cite{combine} The oracle reordering is created by using the permutation of source sentences, which is extracted from the word alignment between the source text and reference.\cite{birch2}

@@

We performed experiments with or without different reordering methods covering the English-to-Chinese translation direction. The reordering methods included our MLT reordering approach and the other reordering approaches with short rules, long rules and tree rules. The system was trained on news text from the LDC corpus and subtitles from TED talks. The development data and test data were both news text from the LDC corpus. The system was a phrase-based SMT system, which used a $6$-gram language model with Knersey-Ney smoothing. Besides the preordering, no lexical reordering or other reordering method in decoding phase was used. The text was translated through a monotone decoder.

@@

Table~\ref{tenw} shows the BLEU scores, absolute improvements of BLEU scores and TER scores for configurations with different reordering methods. The table consist of $2$ sections. the first row of the top section shows results of the baseline, which used no preordering at all. In the following rows of the top section, different types of reordering rules are combined gradually, with each type per row, and the results are showed. For example, the row with \emph{$+$MLT} Rules presents the configuration with all the rule types including MLT rules and all the other rules in the rows above. All improvements are absolute improvements of BLEU scores in comparison to the baseline. Each row with a certain reordering type presents all the different variations of this type and the best score under these configurations are shown. For example, long rules include the left rules and right rules, and the tree rules include the partial rules and recursive application. The baseline used a monotone decoder and no preordering. The row with \emph{oracle reordering} shows the results from the configuration that used the oracle reordering as input. The results of oracle reordering can be used for analyzing the potential of source sentence reordering. In the lower section of the table, different rule types are not combined and the effect of each rule type is shown.

@@

The experiments for Chinese-to-English systems had a similar setup as described in the last section. The parallel data used in the English-to-Chinese system was also used in this experiment by switching the source language and the target language. We only used the LDC data set for training, and no TED data were used in this system. The test data had three English references for evaluating the results instead of one as in the previous system. The data used are summarized in table~\ref{dzhen2}.

@@

Table~\ref{tzhen2} shows the results for configurations with different reordering methods for the Chinese-to-English translation. The table can be interpreted in the same manner as table~\ref{tenw} in the previous section.

@@

We've also found improvements in the sentence structure. Table~\ref{te} shows some translation examples in both translation directions. Sections are separated by double lines in the table. Each section of this table shows one translation example with the source sentence (\emph{source}), translation without using MLT reordering (\emph{no MLT}), translation with MLT reordering (\emph{MLT}) and the reference (\emph{reference}). The translation without using MLT reordering comes from the configuration with highest BLEU score that didn't use MLT reordering. And the translation with MLT reordering comes from the configuration with highest BLEU score that used MLT reordering. From the examples, we can clearly see the improvements in sentence structure.

@@

Source & the dalai lama will go to visit washington this month .\\ \hline
No MLT & \cntext{达赖 喇嘛 将 访问 华盛顿 的 这 一 个 月 。} \\ \hline
MLT & \cntext{达赖 喇嘛 将 本 月 访问 华盛顿 。} \\ \hline
%? add more examples?
Reference & \cntext{达赖 喇嘛 将 在 本 月 前往 华盛顿 访问 。} \\ \hline \hline
Source & \cntext{陈至立 说 , 古巴 是 拉美 和 加勒比 地区 有 重要 影响 的 国家 。}
\\ \hline
No MLT & chen zhili said : cuba is the latin america and the caribbean region has an important influence on the state .
\\ \hline
MLT & chen zhili said : cuba is a country of important influence latin america and the caribbean region .
\\ \hline
Reference & chen zhili said that cuba is a country of great influence in the latin american and caribbean region .
\\ \hline \hline

Source & \cntext{近年 来 , 两 国 教育 交流 日益 密切 , 人员 来往 频繁 。}
\\ \hline
no MLT & in recent years , the two countries education have been increasingly close exchanges and personnel contacts have been frequent .
\\ \hline
MLT & in recent years , the educational exchanges between the two countries have become increasingly frequent , and have had frequent contacts .
\\ \hline
Reference & in recent years , the educational exchange between the two countries has become increasingly close with frequent personnel visits .
\\ \hline

\end{tabular}
\caption{Examples of translations}
\label{te}
\end{table*}

However, by taking a close look at the gap between the scores of oracle reordering and the best scores achieved by MLT reordering, we can also see, there's still potential for improvements of translation between English and Chinese through better reordering methods.

@@

\section{Conclusions}

We've present a new preordering approach for translation between English and Chinese. The algorithm detects and applies reordering rules uses information of word alignment and multiple syntactic levels in the syntax tree to reorder the source sentences. Reordering patterns are detected by checking if the nested tag sequences in subtrees with any number of search levels have clearly new orders in the aligned text in the target language.

>>>>

\documentclass[a4paper]{article}
\usepackage{iwslt14,amssymb,amsmath,epsfig,multirow,spreadtab,array}
\usepackage[normalem]{ulem}
\usepackage{tikz}
\usetikzlibrary{arrows}
\usetikzlibrary{trees}
\usetikzlibrary{shapes}
\usetikzlibrary{chains}
\usetikzlibrary{calc}
\usetikzlibrary{positioning}
\usepackage[T1]{fontenc}
\usepackage[latin1,utf8x]{inputenc}
\usepackage[encapsulated]{CJK}
\newcommand{\cntext}[1]{\begin{CJK}{UTF8}{gbsn}#1\end{CJK}}
\setcounter{page}{1}
\sloppy		% better line breaks
%\ninept
%SM below a registered trademark definition
\def\reg{{\rm\ooalign{\hfil
     \raise.07ex\hbox{\scriptsize R}\hfil\crcr\mathhexbox20D}}}

@@

\title{Rule-Based Preordering on
Multiple Syntactic Levels \\in Statistical Machine Translation}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Please make sure to keep technical paper submissions anonymous  !
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\name{Firstname Lastname}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% If multiple authors, uncomment and edit the lines shown below.       %%
%% Note that each line must be emphasized {\em } by itself.             %%
%% (by Stephen Martucci, author of spconf.sty).                         %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \makeatletter
 \def\name#1{\gdef\@name{#1\\}}
 \makeatother
 \name{{\em Ge Wu, Yuqi Zhang, Alex Waibel}}
%%%%%%%%%%%%%%% End of required multiple authors changes %%%%%%%%%%%%%%%%%

\address{Institute for Anthropomatics \\
Karlsruhe Institute of Technology, Germany \\
{\small \tt utcur@student.kit.edu, yuqi.zhang@kit.edu, alexander.waibel@kit.edu }
}
%
\begin{document}
\maketitle
%
\begin{abstract}
We propose a novel data-driven rule-based preordering approach, which uses the tree information of multiple syntactic levels. This approach extend the tree-based reordering from one level into multiple levels, which has the capability to process more complicated reordering cases. We've conducted experiments in English-to-Chinese and Chinese-to-English translation directions. Our results show that the approach has led to improved translation quality both when it was applied separately or when it was combined with some other reordering approaches. As our reordering approach was used alone, it showed an improvement of $1.61$ in BLEU score in the English-to-Chinese translation direction and an improvement of $2.16$ in BLEU score in the Chinese-to-English translation direction, in comparison with the baseline, which used no word reordering. As our preordering approach were combined with the short rule \cite{short}, long rule \cite{long} and tree rule \cite{tree} based preordering approaches, it showed further improvements of up to $0.43$ in BLEU score in the English-to-Chinese translation direction and further improvements of up to $0.3$ in BLEU score in the Chinese-to-English translation direction. Through the translations that used our preordering approach, we've also found many translation examples with improved syntactic structures.

%syntax tree and word alignment to reorder the words in source sentences before decoding in a phrase-based SMT system between English and Chinese. The preordering algorithm extracts reordering patterns from multiple levels of the syntax trees of the training data and applies the rules on the source sentences in a similar way.

\end{abstract}
%
\section{Introduction}
\label{in}

Word order is a general issue when we want to translate text from one language to another. Different languages normally have different word orders and the difference could be very huge. Among all the languages, Chinese is one language which is very different from English, because they belong to different language families and have long period of separately development. Both languages have a Subject-Verb-Object order, but they also have a lot of differences in word order. Especially sentences in both languages can sometimes have completely different syntactic structures. The differences may involve long-distance or unstructured position changes.

@@

The rest of this paper is organized as follows. Section~\ref{re} presents a review of related work. In Section~\ref{mo}, we point out the problems for translation between English and Chinese and describe the motivation of MLT reordering. In Section~\ref{mu}, we introduce the details of the MLT preordering. In Section~\ref{ex}, we present the experiment results and evaluation. Finally, we conclude this paper in Section~\ref{co}.

\section{Related Work}
\label{re}

@@

In a phrase-based SMT system, there are several possibilities to change the word orders. Words can be reordered during the decoding phase by setting a window, which allows the decoder to choose the next word for translation. Reordering could also be influenced by the language model, because the language model gives probability of how a certain word is likely to follow. Different language model may give different probability, which further influences the decision made by log-linear model. Other ways to change the word orders include using distance based reordering models or lexicalized reordering models \cite{tillmann2004, koehn2005}. The lexicalized reordering model reorders the phrases by using information of how the neighboring phrases change orientations.

@@

However, these approaches which are based on POS tag sequences or syntax trees are mostly designed for languages like German and are not especially adapted for languages like Chinese. As Chinese has very different word orders, a reordering approach, which can further explore the syntactic structure of Chinese and utilize this information for reordering, is desirable.

@@

In this paper, we compare our results to the aforementioned rule-based reordering methods and the oracle reordering to get a better overview.

\section{Motivation}
\label{mo}

@@

Some typical problems of word orders between English and Chinese that we have found are as follows:

\begin{itemize}
\item \emph{Pre-modifier instead of post-modifier}\\ In Chinese people tend to use pre-modifier rather than post-modifier. This involves the position change of adverbials, relative clauses and preposition phrases during translation. Figure~\ref{relative} shows an example of how the position of a relative clause changes
\begin{figure}
\input{relative.tikz}
\caption{Position change of a relative clause}
\label{relative}
\end{figure}
\item \emph{Construction of questions}\\
The two languages have very different ways to construct questions, which raises word order problems for translating questions. Figure~\ref{question} shows word reordering of a question.
\begin{figure}
\input{question.tikz}
\caption{Word reordering of a question}
\label{question}
\end{figure}
\item \emph{Special sentence constructions}\\ For example, \emph{B\^{a}}-construction (\cntext{把字句}) in Chinese and sentence constructions such as \emph{there-be} and inverted negative sentences in English do not have correspondence in the other language in general.
\item \emph{Long distance word position change}\\
Word reordering between English and Chinese often involves word shift cross long distance. For example, following translation shows that a adverbial clause (underlined) is shifted cross long distance when being translated.\bigskip \\
\parbox{0.3705\textwidth}{
I find this very much disturbing \uline{when we are talking about what is going on right and wrong with democracy these days}.\medskip\\
\cntext{\uline{现在，每当我跟别人讨论我们的民主什么是}}
\cntext{\uline{对的，什么是错的}我都为此觉得很无力。}
}
\end{itemize}

Following the analysis above, in order to improve the reordering between English and Chinese, we need reordering method that can handle more complicated, unstructural word order change. Inspired by the ideas of reordering on syntax tree and hierarchical phrases, we created the Multiple-Level-Tree(MLT) reordering, which reorders words based on multiple syntactic levels and can handle long distance word position change and complicated word position change very well.

\section{Multi-Level-Tree Reordering}
\label{mu}

Our preordering method is based on automatically learned reordering rules. Reordering rules show how sentences should be reordered in source language before translation. In our system, the rules are generated by using the word alignment, and syntax tree, all of which are calculated based on the training data. After reordering rules are applied to the source sentences, word lattices are generated. A word lattice contains all the reordering possibilities of a source sentence and is further passed to the decoder for translation. The preordering system could be illustrated as Figure~\ref{prereordering}.

\begin{figure*}
\centering
\tikzset{every picture/.style={scale=0.8}}%
\input{prereordering.tikz}
\caption{Illustration of preordering system}
\label{prereordering}
\end{figure*}

@@

Reordering patterns are based on multiple levels of the syntax tree. Figure~\ref{unstructured} illustrates how the reordering patterns are detected from the syntax tree. In the example, the detection starts from the root node, go downwards three levels and use the nodes in these levels to detect the reordering pattern. These nodes that are used for detecting the reordering pattern are colored gray and have an italic font. Especially, the nodes with dark gray are the lowest ones among such nodes, which represent constituents that are actually reordered through the reordering rules. The leaf nodes in the syntax tree indicate words in the sentence, which has rectangle shape with angular corners.

According to the alignment information with the corresponding Chinese translation, the node labeled with \emph{NP} is moved forward to the first place in the translation and the node labeled with \emph{IN-of} is moved forward to the second place in translation. This reordering cannot be handled with one-level tree-based reordering, but in MLT, from the root node with a search depth of three, the following reordering pattern can be found:\bigskip

\noindent \texttt{NP(\textbf{CD}$_0$ NP(NP(\textbf{JJ}$_1$ \textbf{NNS}$_2$)PP(\textbf{IN}$_3$ \textbf{NP}$_4$)))}\medskip\\
\texttt{-> NP IN CD JJ NNS}\\
\texttt{-> 4 3 0 1 2 (alternative with index)}\bigskip

The POS tags in bold corresponds to nodes with dark gray in Figure~\ref{unstructured}, which presents constituents that are actually reordered. The parentheses in the reordering pattern indicate the corresponding hierarchies in the syntax tree. The reordering can be alternatively represented with index, which is the actual internal representation to avoid ambiguity caused by POS tags with the same name and we use this representation throughout this paper.

\begin{figure}[ht]
\centering
%\tikzset{every picture/.style={scale=0.8}}%
\input{unstructured.tikz}
\caption{Detection of reordering pattern from multiple syntactic levels}
\label{unstructured}
\end{figure}

@@

In order to find as much information for reordering as possible, the algorithm of rule extraction detects the reordering patterns from all nodes in the syntax tree and it goes downwards for any number of hierarchies, until it reaches the lowest hierarchy in the subtrees.

@@

\begin{figure}[ht]
\centering
\input{extract.tikz}
\caption{A phrase with its syntax tree and word alignment for rule extraction}
\label{extract}
\end{figure}

Figure~\ref{extract} shows a phrase to be translated, together with its syntax tree and word alignment of parallel text. In this example, we can find the following reordering patterns:\bigskip

\noindent From node $1$:\\
\texttt{NP(\textbf{NP} \textbf{PP})->1 0} \hfill [1 level]\hphantom{s}\\
\texttt{NP(NP(\textbf{JJ} \textbf{NNS})PP(\textbf{IN} \textbf{NP}))->3 2 0 1} \hfill [2 levels]\\
\texttt{NP(NP(\textbf{JJ} \textbf{NNS})PP(\textbf{IN} NP(\textbf{JJ} \textbf{NNS})))-> \\ 3 4 2 0 1} \hfill [3 levels]\medskip

\noindent From node $3$:\\
\texttt{PP(\textbf{IN} \textbf{NP})->1 0} \hfill [1 level]\hphantom{s}\\
\texttt{PP(\textbf{IN} NP(\textbf{JJ} \textbf{NNS}))->1 2 0} \hfill [2 levels]\bigskip

@@

The syntax tree is traversed by DFS as the same in rule extraction. But from the root of each subtree, it has scanned with depth limit from its maximal levels, i.e. its depth, to $1$. As it turns out that any rule can be applied for a subtree at some level, a new path for this reordering will be added to the word lattice for decoding. As long as rules can be applied on a subtree for a certain depth, the search for rule application on this subtree stops, and the search on the next subtree continues.

The reason for this is to prevent duplicate reorderings due to application of nested rules, which have overlapped effect with each other. These rules are normally patterns that are generated on the same subtree, but with different number of levels, which has different generalization effect on the same range of words in the text. For example, the following patterns can be detected from the syntax tree in Figure~\ref{extract}:\bigskip

\noindent \texttt{PP(\textbf{IN} \textbf{NP}) -> 1 0}\\
\texttt{PP(\textbf{IN} NP(\textbf{JJ} \textbf{NNS})) -> 1 2 0}\bigskip

Both patterns are detected from the same node, but the second pattern is detected by retrieving the nodes one level deeper and it is more concrete. So the first pattern can be seen as a generalization of the second pattern. Whenever a rule of the second pattern can be applied, a rule of the first pattern can be applied too. Because subtrees are checked from the highest number of levels in rule application, the more concrete rule is applied first. Since the more concrete rule fits the detected pattern better and contains more details of reordering, so it may be more suitable for rule application. In this example, the second rule is applied rather than the first rule.

@@

In order to further explore the probability of improvement, the MLT reordering rules can be combined with other types of reordering rules to achieve further improvements. This is done by training the different types of rules separately and applying them on the monotone path of the sentence independently. All the generated different paths are compressed in the word lattice.

\section{Experiments}
\label{ex}

We have conducted experiments on both English-to-Chinese and Chinese-to-English translation directions to get a better overview of the MLT reordering's effect. In the experiments, the MLT reordering rules are also combined with other reordering rules that are introduced before, in order to show the improvement achieved by our approach.

In order to evaluate the potential of word reordering, we also used oracle reordering in the experiments. Oracle reordering is considered to be an optimally reordered sentence as input to the SMT system and do not allow additional reordering during decoding \cite{combine}. The oracle reordering is created by using the permutation of source sentences, which is extracted from the word alignment between the source text and reference \cite{birch2}.

@@

We performed experiments with and without different reordering methods covering the English-to-Chinese translation direction. The reordering methods included our MLT reordering approach and the other reordering approaches with short rules, long rules and tree rules. The system was trained on news text from the LDC corpus and subtitles from TED talks. The development data and test data were both news text from the LDC corpus. The system was a phrase-based SMT system, which used a $6$-gram language model with Knersey-Ney smoothing. Besides the preordering, no lexical reordering or other reordering method in decoding phase was used. The text was translated through a monotone decoder.

@@

Table~\ref{tenw} shows the BLEU scores, absolute improvements of BLEU scores and TER scores for configurations with different reordering methods. The table consists of $2$ sections. the first row of the top section shows results of the baseline, which used no preordering at all. In the following rows of the top section, different types of reordering rules are combined gradually, with each type per row. For example, the row with \emph{$+$MLT Rules} presents the configuration with all the rule types including MLT rules and all the other rules in the rows above. All improvements are absolute improvements of BLEU scores in comparison to the baseline. Each row with a certain reordering type presents all the different variations of this type and the best score under these configurations is shown. For example, long rules include the left rules and right rules, and the tree rules include the partial rules and recursive application. The baseline used a monotone decoder and no preordering. The row with \emph{oracle reordering} shows the results from the configuration that used the oracle reordering as input. The results of oracle reordering can be used for analyzing the potential of source sentence reordering. In the lower section of the table, different rule types are not combined and the effect of each rule type is shown.

@@

The experiments for Chinese-to-English systems have a similar setup as described in the last section. The parallel data used in the English-to-Chinese system was also used in this experiment by switching the source language and the target language. We only used the LDC data set for training, and no TED data were used in this system. The test data had three English references for evaluating the results instead of one as in the previous system. The data used are summarized in Table~\ref{dzhen2}.

@@

Table~\ref{tzhen2} shows the results for configurations with different reordering methods for the Chinese-to-English translation. The table can be interpreted in the same manner as Table~\ref{tenw} in the previous section.

@@

We have also found improvements in the sentence structure. Table~\ref{t1} and Table~\ref{t2} shows some translation examples in both translation directions. Sections are separated by double lines in the table. Each section of this table shows one translation example with the source sentence (\emph{source}), translation without using MLT reordering (\emph{no MLT}), translation with MLT reordering (\emph{MLT}) and the reference (\emph{reference}). \emph{Glossary} for the source sentences in translation examples from English to Chinese is also added for understanding. The translation without using MLT reordering comes from the configuration with highest BLEU score that didn't use MLT reordering. And the translation with MLT reordering comes from the configuration with highest BLEU score that used MLT reordering. From the examples, we can clearly see the improvements in sentence structure.

% Each word or group of hyphenated words in the glossary corresponds a Chinese character or a group of Chinese characters that are not separated with space in the source sentence with the same order.

\begin{table*}
\centering
\begin{tabular}{|l|m{0.7\textwidth}|} \hline
Source & \cntext{陈至立 说 , 古巴 是 拉美 和 加勒比 地区 有 重要 影响 的 国家 。}
\\ \hline
Glossary & chen-zhili said , cuba is latin-american and caribbean region has great influence of country .
\\ \hline
No MLT & chen zhili said : cuba is the latin america and the caribbean region has an important influence on the state .
\\ \hline
MLT & chen zhili said : cuba is a country of important influence latin america and the caribbean region .
\\ \hline
Reference & chen zhili said that cuba is a country of great influence in the latin american and caribbean region .
\\ \hline \hline

Source & \cntext{近年 来 , 两 国 教育 交流 日益 密切 , 人员 来往 频繁 。}
\\ \hline
Glossary & recent-years in, two countries educational exchange increasingly close , personnel visits frequent .
\\ \hline
no MLT & in recent years , the two countries education have been increasingly close exchanges and personnel contacts have been frequent .
\\ \hline
MLT & in recent years , the educational exchanges between the two countries have become increasingly frequent , and have had frequent contacts .
\\ \hline
Reference & in recent years , the educational exchange between the two countries has become increasingly close with frequent personnel visits .
\\ \hline
\end{tabular}
\caption{Examples of translations from Chinese to English}
\label{t1}
\end{table*}

@@

Source & the dalai lama will go to visit washington this month .\\ \hline
No MLT & \cntext{达赖 喇嘛 将 访问 华盛顿 的 这 一 个 月 。} \\ \hline
MLT & \cntext{达赖 喇嘛 将 本 月 访问 华盛顿 。} \\ \hline
%? add more examples?
Reference & \cntext{达赖 喇嘛 将 在 本 月 前往 华盛顿 访问 。} \\ \hline
\end{tabular}
\caption{Examples of translations from English to Chinese}
\label{t2}
\end{table*}

@@

\section{Conclusions}
\label{co}

We have presented a new preordering approach for translation
between English and Chinese. The algorithm detects and applies reordering rules by using information of multiple syntactic levels in the syntax tree and word alignment to reorder the source sentences. Reordering patterns are detected by checking if the nested tag sequences in subtrees with any number of search levels have clearly new orders in the aligned text in the target language.

We have conducted experiments in both translation directions between English and Chinese with different SMT configurations. From the results we can see the BLEU scores were improved no matter when we applied the SMT reordering method to the baseline directly or when we combined it with the other reordering methods, i.e. short rules, long rules and tree rules based reordering methods.
